\chapter{Article 3: Resource Classification from Version Control System Logs}
%\chapter{Discovering the Organizational Perspective}
\label{chap:resource-classfication}

{\bfseries \Large Authors: \medskip}

Kushal Agrawal,
Michael Aschauer,
Thomas Thonhofer,
Nico Tomsich, 
Saimir Bala, and 
Andreas~Rogge-Solti \hfill

\bigskip

{\noindent\bfseries \Large Published in: \medskip}

20th {IEEE} International Enterprise Distributed Object Computing
Workshop, {EDOC} Workshops 2016, Vienna, Austria, September 5-9, 2016

\bigskip

{\noindent\bfseries \Large Abstract: \medskip}


Collaboration in business processes and projects requires a division of responsibilities among the participants. Version control systems allow us to collect profiles of the participants that hint at participants' roles in the collaborative work. The goal of this paper is to automatically classify participants into the roles they fulfill in the collaboration. Two approaches are proposed and compared in this paper. The first approach finds classes of users by applying k-means clustering to users based on attributes calculated for them. The classes identified by the clustering are then used to build a decision tree classification model. The second approach classifies individual commits based on commit messages and file types. The distribution of commit types is used for creating a decision tree classification model. The two approaches are implemented and tested against three real datasets, one from academia and two from industry. Our classification covers 86\% percent of the total commits. The results are evaluated with actual role information that was manually collected from the teams responsible for the analyzed repositories.

\pagebreak


%\todo[inline]{Polish the chapter}


\input{ResourceClassification/sections/Introduction}
\input{ResourceClassification/sections/Background}
\input{ResourceClassification/sections/Concept}
\input{ResourceClassification/sections/Implementation}

\section{Discussion}

Throughout this project we discovered a number of research directions to follow, methods to use and features to analyze, which made our research much more exploratory than originally planned. For our research we picked from a vast selection of potential methods and tools, a small set which seemed promising to us. %Considering the experimental character of the project and the fact that we did not really have an idea of what our findings would look like up front, we achieved quite good results. 
The user-based approach, which was our initial plan, provided a useful insight into the data but fell short of providing generally applicable results in terms of a classification algorithm. The commit-based approach seems more valuable to us, due to its results and the much larger spectrum of information it delivers for analyzing users. The commit classification method seems quite robust and could with some extensions and refinements become a useful tool for analyzing arbitrary VCS repositories. The user profiles created based on this classification are definitely an interesting source of information and building a classification model on top of them has proven to be a legitimate approach. 

Our first research question (R1) can be answered with: Yes, it is possible to classify users based on the information usually found in VCS logs. Our results show that at least for some typical roles in software development there can be enough information in commit messages and file types, to make certain statements about the users who created them. However this is not true for all users, especially when their commit styles and frequencies are very different.

The other two questions (R2 and R3) are more difficult to answer. The classes to be found can differ between repositories. From our experience, developers can be distinguished from other roles quite easily due to high numbers of commits and/or modified files as well as the usage of certain words and phrases. Testers can also be set apart in most cases, mainly through key words. Looking at the file types of added, modified and deleted files also reveals users in the field of web development. Beyond that there are certainly more differences and classes to be found but they are less obvious and require more research.

What do these results mean for the practical use of resource classification using VCS logs as discussed in \Cref{sec:problemDescription}? One clear finding of our work is that any project team trying to use this kind of classification, should specify and enforce some guidelines for how VCS should be used. They should at least make the usage of commit messages mandatory, ideally also define a structure and vocabulary for them. While the models we created are probably not applicable for an arbitrary team or project, a team could follow our approach to create their own individual model using specific knowledge. For a more generally applicable classification model, additional research and improvements to our solution would be necessary. 

While our algorithms and models work well for describing our three data sets, it still needs to be tested for other repositories, especially from other domains. This would require testing with more data sets, more verification information and more manual analysis. While fetching more VCS repositories and running them through our algorithms could be done in a short amount of time, obtaining the necessary role information is challenging and the manual analysis time consuming.

A current limitation of analyzing VCS logs is that commit messages are used differently between users, repositories and teams. We have to assume that it is challenging to find one classification model that produces valid results for any VCS repository, because there are ambiguities in what certain statements mean. In the worst case, users omit commit messages, which renders classification based on text impossible.

%\section{Summary}\label{sec:conclusion}
\input{ResourceClassification/sections/Conclusion}
