%\chapter{Further Perspectives for Monitoring the Software Process}

\chapter{Article 5: Software Process Evaluation from User Perceptions and Log Data}


{\bfseries \Large Authors: \medskip}

\noindent Damjan Vavpotič,
Saimir Bala,
Jan Mendling and,
Tomaž Hovelja \hfill

\bigskip

{\noindent\bfseries \Large Published in: \medskip}

\noindent Journal of Software: Evolution and Process

\bigskip

{\noindent\bfseries \Large Abstract: \medskip}


\noindent \textbf{Purpose} – Companies often claim to follow specific software development methodologies (SDM) when performing their software development process. These methodologies are often supported by dedicated tools that keep track of work activities carried out by developers. The purpose of this paper is to provide a novel approach that integrates analytical insights from both the perceptions of SDM stakeholders and software development tools logs to provide SDM improvement recommendations. \\

\noindent
\textbf{Design/methodology/approach} – This paper develops a new process improvement approach that combines two significantly different sources of data on the same phenomenon. First, it uses a questionnaire to gather software development stakeholder SDM perceptions (managers, developers). Second, it leverages process mining to analyse software development tools logs to obtain additional information on software development activities. Finally, it develops recommendations based on concurrent analysis of both sources. \\

\noindent
\textbf{Findings} – Our novel process improvement approach is evaluated in three directions: does the presented approach (RQ1) enable managers to gain additional insights into employees’ performance, (RQ2) deliver additional insights into project performance, and (RQ3) enable development of additional SDM improvement recommendations. We find that integrated analysis of software development perception data and software development tools logs opens new possibilities to more precisely identify and improve specific SDM elements.\\

\noindent
\textbf{Research limitations} – The evaluation of our novel process improvement approach follows a single case study design. Our approach can only be used in enterprises in which software development tools logs are available. The study should be repeated in different cultural settings. \\

\noindent
\textbf{Practical implications} – We practically show how concurrently analysing data about developer SDM perceptions and event log data from software development tools enables management to gain additional insights in the software development process regarding the performance of individual developers. \\

\noindent
\textbf{Originality/value} – The main theoretical contribution of our paper is a novel process improvement approach that effectively integrates data from management and developer perspectives and software development tools logs. \\

\noindent
\textbf{Keywords} – software development methodology, evaluation, stakeholder perceptions, software development tool logs, case study

\pagebreak


% Management/Monitoring Perspective and How these techniques contribute
% Trace data 

% 
%
%
%Make the title shorter.
%
%\todo[inline]{Polish the chapter}
%\todo[inline]{Decide with Jan what to do here. Keep/Kill?}
%\todo[inline]{Elaborate more on the text. Motivate this section by including the social science point of view. The previous chapters are belong more to software engineering.}

%The techniques presented in this dissertation fall under the broader category of analyses methods based on trace data. When used individually they provide specific insights about determinate perspective of a process. These methods can also be combined with one another in order to achieve a more complete picture about how the process performs under the different perspectives. Beyond measuring performance of processes, trace-data methods can be exploited in combination with other methods to aid theory development~\citep{DBLP:journals/isr/BerenteSS19}.
%
%This chapter presents two more approaches that leverage trace data to gather insights on the software development. \Cref{sec:combining-questionnaier-and-trace} explores a mixed-method approach to identify relevant software development elements. It combines a subjective perspective from questionnaires and data analyses from event logs. As a result, insights are gathered on what elements both considered important and are actually used. \Cref{sec:conversational-structure} presents an approach that leverages speech act theory to understand how conversation about software development unfold. As a result, it is possible to observe differences in the process of successful collaboration. 

%\input{chap/ch6DamjanUsefulnes}
\input{chap/art5DamjanFull}

%\input{chap/ch7Aron}