\section{Running example}\label{sec:running_example}

In this paper we will use an example scenario taken from the SHAPE project case. The project history starts when people get together and begin to work on a proposal. This phase involves the definition of the project goals and milestones, as well as the work-packages that address them. After the cost plan and resources are estimated, a kickoff meeting is planned and resources are assigned to work-packages. At this point in time, a hierarchical organization of the repository, that reflects project work packages, is defined. As the project goes on following regular meetings and milestone deadlines, resources that are assigned to work packages, contribute by adding pieces of work over time. 

For sake of simplicity, in our running example we show the log data of a specific time period and task, namely \emph{example}. Such task, was assigned to resource X after a meeting, when a part of project members asked for a toy example to better understand the problem domain. During his activity, resource X was supervised by his direct responsible Y and his work was then reviewed, after another meeting, by Z. 

An extract from the log concerning this period, that shows some crucial steps in the \emph{example} task, is reported on table \ref{tab:example}. Comments were left out from this representation, since at the current state we don't delve into them.   

\begin{table}\scriptsize
\caption{Excerpt from VCS log data for the referenced time period }
\label{tab:example}

\renewcommand{\arraystretch}{1.3}
\centering
%\fontfamily{phv}\fontseries{mc}\selectfont
\begin{tabular}{m{.8cm} m{1.5cm} m{3cm} p{5.8cm}}
%\hline\noalign{\smallskip
%\hline
\textbf{CID}	 & \textbf{Resource} & \textbf{Date} & \textbf{List of changes} \\
%\noalign{\smallskip}
\hline 
%\noalign{\smallskip}
\hline

\multirow{2}{*}{204} & \multirow{2}{*}{Y} & \multirow{2}{*}{2014-11-12~11:57:46} & A /example \\
& & & A \slash example\slash SHAPE\slash\-ToyStation\-Example.docx \\ \hline %\hdashline

%\multirow{2}{*}{205} & \multirow{2}{*}{X} & \multirow{2}{*}{2014-11-14 16:26:23} & A /example/ToyStation.bpmn \\
%& & & A /example/ToyStation.png \\ \hline %\hdashline

\ldots & \ldots & \ldots & \ldots \\ \hline

\multirow{2}{*}{206} & \multirow{2}{*}{X} & \multirow{2}{*}{2014-11-14 16:34:07} & M /example/ToyStation.bpmn\\
& & & M /example/ToyStation.png \\ \hline %\hdashline

%\multirow{7}{*}{207} & \multirow{7}{*}{X} & \multirow{7}{*}{2014-11-27 14:18:59} & M /example/SHAPE-ToyStation-Example.docx \\
%& & & D /example/ToyStation.bpmn \\
%& & & M /example/ToyStation.png \\
%& & & A /example/ToyStation\_0Loop.bpmn \\
%& & & A /example/ToyStation\_1Loop.bpmn \\
%& & & A /example/ToyStation\_nLoop.bpmn \\
%& & & A/example/ToyStation\_old.bpmn\\ \hline

\ldots & \ldots & \ldots & \ldots \\ \hline

222 & W & 2014-12-15 13:49:11 & D /example/Download \\ \hline %\hdashline

223 & W & 2015-01-08 16:06:41 & A /example/Download2\\ \hline %\hdashline

\multirow{2}{*}{225} & \multirow{2}{*}{X} & \multirow{2}{*}{2015-01-13 11:47:09} & M /example/ToyStation\_0Loop.bpmn\\
& & & M /example/ToyStation\_nLoop.bpmn \\ \hline %\hdashline

\multirow{2}{*}{226} & \multirow{2}{*}{Z} & \multirow{2}{*}{2015-01-23 16:50:29} & A /example/ToyStation\_0Loop.pdf\\
& & & A /example/ToyStation-feedbackZ.pdf\\ \hline %\hdashline
\end{tabular}\\ \hfill

\end{table}
%\vfill
